---
title: "p8105_hw1_jys2136"
author: "Jocelyn Sun"
date: "9/29/2021"
output: github_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

# Problem 1

## Creating a data frame 

```{r problem_1}

set.seed(20)

prob1_df = tibble(
  random_samp = rnorm(10),
  vec_pos = random_samp > 0,
  vec_char = c("apple", "orange", "strawberry", "banana", "mango", "grape", "plum", "persimmon", "grapefruit", "fig"),
  vec_fac = factor(c("April", "May", "June", "April", "May", "June", "April", "May", "June", "April"))
)

mean(pull(prob1_df, random_samp))
mean(pull(prob1_df, vec_pos))
mean(pull(prob1_df, vec_char))
mean(pull(prob1_df, vec_fac))

```

For the vectors that include characters ("vec_fac" and "vec_char"), I receive a warning message informing that the argument is not numeric or logical. R produces "NA" as a result. 

I am able to receive results for "random_samp" and "vec_pos" because they contain numerical values. This exercise shows that R does not compute the mean for non-numeric or character values.

## Applying the as.numeric function

```{r problem_1_convert, eval = FALSE}

as.numeric(pull(prob1_df, (vec_pos))
as.numeric(pull(prob1_df, vec_char))
as.numeric(pull(prob1_df, vec_fac))

```

I am able to convert "vec_fac" and "vec_pos", but not "vec_char" 

The as.numeric argument recognizes that logic vector has two possible outcomes "TRUE" or "FALSE" and converts this into binary "0" or "1"

The as.numeric argument recognizes that the factor vector has 3 distinct levels and converts each level to a number: "1", "2", or "3", respective of the original 
"levels"


three different levels, thus providing a proverly converted numbers.  However, R was unable to take the mean of the factor variable when the levels were non-numeric. 


# Problem 2
```{r problem_2}

data("penguins", package = "palmerpenguins")

skimr::skim(penguins)

view(penguins)

nrow(penguins)
ncol(penguins)

mean((pull(penguins, flipper_length_mm)), na.rm = TRUE)

```

From `skimr::skim(penguins)` we see that there are 3 factor variables and 5 numeric variables. In total, this data set has 8 variables. 

`nrow(penguins)` shows that we have 344 rows.
`ncol(penguins)` shows that we have 8 columns.

The mean flipper length is `r mean((pull(penguins, flipper_length_mm)), na.rm = TRUE)` mm, which is given by `mean((pull(penguins, flipper_length_mm)), na.rm = TRUE)` and in the data overview  provided by `skimr::skim(penguins)`.


## Creating a scatterplot
```{r problem_2_scatter}

ggplot(penguins, 
       aes(x = bill_length_mm, 
           y = flipper_length_mm, 
           color = species)) + geom_point()

```


```{r scatter_save}

ggsave("p8105_hw1_jys2136.pdf", height = 4, width = 6)

```
