---
title: "p8105_hw1_jys2136"
author: "Jocelyn Sun"
date: "9/29/2021"
output: github_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

# Problem 1

## Creating a data frame 

```{r problem_1, warning = FALSE}

set.seed(20)

prob1_df = tibble(
  random_samp = rnorm(10),
  vec_pos = random_samp > 0,
  vec_char = c("apple", "orange", "strawberry", "banana", "mango", "grape", "plum", "persimmon", "grapefruit", "fig"),
  vec_fac = factor(c("April", "May", "June", "April", "May", "June", "April", "May", "June", "April"))
)

mean(pull(prob1_df, random_samp))
mean(pull(prob1_df, vec_pos))
mean(pull(prob1_df, vec_char))
mean(pull(prob1_df, vec_fac))

```

For the vectors that include characters (`vec_fac` and `vec_char`), I receive a warning message informing that the argument is not numeric or logical. R produces `NA` in the output and the following warning message:

```
Warning message:
In mean.default(pull(<data>, <vector>)): argument is not numeric or logical: returning NA
```

I am able to receive results for `random_samp` and `vec_pos` because they contain numerical values. This exercise shows that R does not compute the mean for non-numeric or character values.

## Converting to numeric data

```{r problem_1_convert, eval = FALSE}

as.numeric(pull(prob1_df, (vec_pos))
as.numeric(pull(prob1_df, vec_char))
as.numeric(pull(prob1_df, vec_fac))

```

I am able to convert `vec_fac` and `vec_pos`, but not `vec_char` 

The as.numeric argument recognizes that the factor vector `vec_fac` has 3 distinct levels and converts each level to a number: `1`, `2`, or `3`, respective of the original "levels", which explains why we could take the mean.

The as.numeric argument recognizes that logic vector `vec_pos` has two possible outcomes `TRUE` or `FALSE` and converts this into binary `0` or `1`, which explains why we could take the mean.

The character vector `vec_char` was not converted numerically. The output produced as series of `NA` because the values in this vector cannot be converted to numbers, which explains why we could not take the mean.

# Problem 2
```{r problem_2}

data("penguins", package = "palmerpenguins")

```

## Data Overview for Dataset Penguins
```{r problem_2_data_overview}

skimr::skim(penguins)

nrow(penguins)
ncol(penguins)

```

From `skimr::skim(penguins)` we see that there are 3 factor variables and 5 numeric variables. In total, this data set has 8 variables. 

`nrow(penguins)` shows that we have 344 rows.
`ncol(penguins)` shows that we have 8 columns.

## Calculating mean flipper length

```{r problem_2_data_mean}

mean((pull(penguins, flipper_length_mm)), na.rm = TRUE)

```

The mean flipper length is `r mean((pull(penguins, flipper_length_mm)), na.rm = TRUE)` mm, which is given by `mean((pull(penguins, flipper_length_mm)), na.rm = TRUE)` and in the data overview  provided by `skimr::skim(penguins)`.


## Scatterplot (flipper length vs bill length)
```{r problem_2_scatter, warning = FALSE}

ggplot(penguins, 
       aes(x = bill_length_mm, 
           y = flipper_length_mm, 
           color = species)) + geom_point()

```

## Saving a scatterplot
```{r scatter_save, warning =  FALSE}

ggsave("p8105_hw1_jys2136.pdf", height = 4, width = 6)

```
