---
title: "p8105_hw1_jys2136"
author: "Jocelyn Sun"
date: "9/29/2021"
output: github_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

# Problem 1

## Creating a data frame 

```{r problem_1}

set.seed(20)

prob1_df = tibble(
  random_samp = rnorm(10),
  vec_pos = random_samp > 0,
  vec_char = c("apple", "orange", "strawberry", "banana", "mango", "grape", "plum", "persimmon", "grapefruit", "fig"),
  vec_fac = factor(c("April", "May", "June", "April", "May", "June", "April", "May", "June", "April"))
)

mean(pull(prob1_df, random_samp))
mean(pull(prob1_df, vec_pos))
mean(pull(prob1_df, vec_char))
mean(pull(prob1_df, vec_fac))

```

For the vectors that include characters ("vec_fac" and "vec_char"), I receive a warning message informing that the argument is not numeric or logical. R produces "NA" as a result. I am able to receive results for "random_samp" and "vec_pos" because they contain numerical values. This exercise shows that R does not compute the mean for non-numeric or character values.

## Applying the as.numeric function

```{r problem_1_convert, eval = FALSE}

transform(prob1_df, trans_pos = as.numeric(vec_pos))
transform(prob1_df, trans_char = as.numeric(vec_char))
transform(prob1_df, trans_fac = as.numeric(vec_fac))

as.numeric(pull(prob1_df, vec_pos))
as.numeric(pull(prob1_df, vec_char))
as.numberic(pull(prob1_df, vec_fac))

```

I am able to convert "vec_fac" and "vec_pos", but not "vec_char". The factor variable recognizes that my variable has three different levels, hence why I am able to convert "vec_fac" to a numeric variable. However, R was unable to take the mean of the factor variable when the levels were non-numeric. 


# Problem 2
```{r problem_2}

data("penguins", package = "palmerpenguins")

view(penguins)



number_rows = nrow(penguins)
number_columns = ncol(penguins)
number_rows
number_columns


mean(pull(penguins, flipper_length_mm))


```


```{r problem_2_scatter}

ggplot(penguins, 
       aes(x = bill_length_mm, 
           y = flipper_length_mm, 
           color = species)) + geom_point()

```


